---
title: "Stroke_Prediction"
author: "Marriah Lewis and Glory Onyeugbo"
date: "5/10/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
Loading the libraries we a going to use on this project.

```{r}
library(plyr)
library(dplyr)
library(grid)
library(tm)
library(RColorBrewer)
library(ggplot2)
library(reshape)
library(corrplot)
library(forcats)
```

# Reading the Stroke Data

Used the read.csv function to call in the stroke data downloaded from Kaggle.

```{r}
# Set the working directory.
setwd("C:\\Users\\17708\\OneDrive\\Desktop\\IST 707 Project")
# Load the dataset.
stroke.df <- read.csv("Stroke Data.csv")
# Print the first 5 instances of the data.
head(stroke.df[1:5])
```

Read the structure.

```{r}
str(stroke.df) 
#5110 obs. of  12 variables
```

# Data Cleaning
```{r}
# Assign the data frame to a new data set as to not comprimise the integrity of the original
stroke_pred <- stroke.df

# remove the id and bmi columns since they have no importance to the ultimate goal of this project.
stroke_pred <- stroke_pred[,-c(1,10)]

colnames(stroke_pred)
# the bmi columns has been reomoved
```
Remove Missing Values and Duplicates

```{r}
#check for missing values 
(sum(is.na(stroke_pred)))

#check for duplicates 
nrow(stroke_pred[duplicated(stroke_pred),]) #there are no duplicates 
#Just to make sure there are no duplicates 
stroke_prediction <- stroke_pred[!duplicated(stroke_pred),]
```

Data Conversions
```{r}
#check for missing values 
# Convert nominal values to factors (hypertension, heart_disease, stroke)
stroke_prediction$gender          <- factor(stroke_prediction$gender)

stroke_prediction$hypertension    <- factor(stroke_prediction$hypertension)

stroke_prediction$heart_disease   <- factor(stroke_prediction$heart_disease)

stroke_prediction$ever_married    <- factor(stroke_prediction$ever_married)

stroke_prediction$work_type       <- factor(stroke_prediction$work_type)

stroke_prediction$Residence_type  <- as.factor(stroke_prediction$Residence_type)

stroke_prediction$smoking_status  <- factor(stroke_prediction$smoking_status)

stroke_prediction$stroke          <- factor(stroke_prediction$stroke)

# Discetize Age
(youngest_stroke <- min(stroke_prediction$age))
(oldest_stroke  <- max(stroke_prediction$age))
stroke_prediction$age_group <- cut(stroke_prediction$age, breaks = c(0,10,20,30,40,50,60,70,80,90),
                    labels=c("child","teens","twenties","thirties","forties","fifties","sixties", "seventies", "eighties"))

# Rename residence colomn name to fit the structure.
names(stroke_prediction)[names(stroke_prediction) == "Residence_type"] <- "residence_type"

# View The Data Structure
str(stroke_prediction)
```
Create a barplot
```{r}
b <- table(stroke_prediction$age, stroke_prediction$stroke)
b
barplot(b, main = "Stroke Risk across All Ages", beside = TRUE, col= brewer.pal(7, "Spectral"),legend.text = rownames((b)),args.legend=list(x="topright",bty="s"))
```

# EDA
```{r}
# How many people have actually had a stroke?
sum(stroke_prediction$stroke == 1)
sum(stroke_prediction$stroke == 0)

# What are the ages of people who suffered
tapply(stroke_prediction$stroke == 1, stroke_prediction$age_group, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$age_group, sum)

# Which gender suffered from strokes the most.
tapply(stroke_prediction$stroke == 1, stroke_prediction$gender, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$gender, sum)
# Majority are females

# Do strokes tend to happen to people with a prior heart condition.
tapply(stroke_prediction$stroke == 1, stroke_prediction$heart_disease, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$heart_disease, sum)
# Majority of people who suffered from a stroke had no prior heart disease.

# Does marital status have anything to do with strokes.
tapply(stroke_prediction$stroke == 1, stroke_prediction$ever_married, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$ever_married, sum)
# An overwhelming amount of people who has a stroke were married

# Does marital status have anything to do with strokes.
tapply(stroke_prediction$stroke == 1, stroke_prediction$hypertension, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$hypertension, sum)
# Most of them did not have hypertension.

# iS the likehood of stroke impacted by smoking status
tapply(stroke_prediction$stroke == 1, stroke_prediction$smoking_status, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$smoking_status, sum)
# There are more non-smokers and former smokers who had strokes that people that currently smoke and a couple oof unkowns. 

# Residence
# Does marital status have anything to do with strokes.
tapply(stroke_prediction$stroke == 1, stroke_prediction$residence_type, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$residence_type, sum)
# Urban is more than rural

# Work Type
tapply(stroke_prediction$stroke == 1, stroke_prediction$work_type, sum)
tapply(stroke_prediction$stroke == 0, stroke_prediction$work_type, sum)

# Which gender tend to have the most cases of heart disease?
tapply(stroke_prediction$heart_disease == 1, stroke_prediction$gender, sum)

# Which age group has the most cases of heart disease?
tapply(stroke_prediction$heart_disease == 1, stroke_prediction$age_group, sum)

# Which gender smoke the most?
tapply(stroke_prediction$smoking_status == "smokes", stroke_prediction$gender, sum)

# Which gender had the most former smokers.
tapply(stroke_prediction$smoking_status == "formerly smoked", stroke_prediction$gender, sum)

```



Visualizations
```{r}
age_count <- tally(group_by(stroke_prediction, age_group, stroke))

stroke_count <- count(stroke_prediction, stroke)
ggplot(stroke_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = stroke, fill = stroke)) + labs(x = 'Stroke', y = 'Count') + ggtitle('Count by Stroke')

#########################################################################################
age_group_count <- count(stroke_prediction, age_group)
ggplot(age_group_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = age_group, fill = age_group)) + labs(x = 'Age Group', y = 'Count') + ggtitle('Count by Age Group')

#########################################################################################
gender_count <- count(stroke_prediction, gender)
ggplot(gender_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = gender, fill = gender)) + labs(x = 'Gender', y = 'Count') + ggtitle('Count by Gender')

#########################################################################################
hypertension_count <- count(stroke_prediction, hypertension)
ggplot(hypertension_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = hypertension, fill = hypertension)) + labs(x = 'Hypertension', y = 'Count') + ggtitle('Count by Hypertension')

#########################################################################################
hd_count <- count(stroke_prediction, heart_disease)
ggplot(hd_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = heart_disease, fill = heart_disease)) + labs(x = 'Heart Disease', y = 'Count') + ggtitle('Count by Heart Disease')

#########################################################################################
wt_count <- count(stroke_prediction, work_type)
ggplot(wt_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = work_type, fill = work_type)) + labs(x = 'Work Type', y = 'Count') + ggtitle('Count by Work Type')

#########################################################################################
marriage_count <- count(stroke_prediction, ever_married)
ggplot(marriage_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = ever_married, fill = ever_married)) + labs(x = 'Married', y = 'Count') + ggtitle('Count by Marriage')

#########################################################################################
ss_count <- count(stroke_prediction, smoking_status)
ggplot(ss_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = smoking_status, fill = smoking_status)) + labs(x = 'Smoking Status', y = 'Count') + ggtitle('Count by Smoking Status')

#########################################################################################
residence_count <- count(stroke_prediction, residence_type)
ggplot(residence_count, aes(y = n)) + geom_bar(stat = "identity", aes(x = residence_type, fill = residence_type)) + labs(x = 'Residence', y = 'Count') + ggtitle('Count by Residence')
```


Visualization
```{r}

#########################################################################################
ggplot(stroke_prediction, aes(x = gender, y = stroke, fill = gender)) + geom_bar(stat = 'identity') + facet_wrap(~stroke) + labs(x = 'Gender', y = 'Count') + ggtitle('Stroke Count by Gender')

#########################################################################################
ggplot(stroke_prediction, aes(x = age_group, y = stroke, fill = age_group)) + geom_bar(stat = 'identity') + facet_wrap(~stroke) + labs(x = 'Age Group', y = 'Count') + ggtitle('Stroke Count by Age') + theme(axis.text.x = element_text(angle = 90))

#########################################################################################
ggplot(stroke_prediction, aes(x = smoking_status, y = stroke, fill = smoking_status)) + geom_bar(stat = 'identity') + facet_wrap(~stroke) + labs(x = 'Smoking Status', y = 'Count') + ggtitle('Stroke Count by Smoking Status') + theme(axis.text.x = element_text(angle = 90))

#########################################################################################
ggplot(stroke_prediction, aes(x = hypertension, y = stroke, fill = hypertension)) + geom_bar(stat = 'identity') + facet_wrap(~stroke) + labs(x = 'Hypertension', y = 'Count') + ggtitle('Stroke Count by Hypertension')

#########################################################################################
ggplot(stroke_prediction, aes(x = heart_disease, y = stroke, fill = heart_disease)) + geom_bar(stat = 'identity') + facet_wrap(~stroke) + labs(x = 'Heart Disease', y = 'Count') + ggtitle('Stroke Count by Heart Disease')

#########################################################################################
ggplot(stroke_prediction, aes(x = work_type, y = stroke, fill = work_type)) + geom_bar(stat = 'identity') + facet_wrap(~stroke) + labs(x = 'Work Type', y = 'Count') + ggtitle('Stroke Count by Work Type') + theme(axis.text.x = element_text(angle = 90))


#########################################################################################
ggplot(stroke_prediction, aes(x = age, y = avg_glucose_level, color = stroke)) + geom_point() + labs(x = 'Age', y = 'Avg Glucose') + ggtitle('Average Glucose by Age')

#########################################################################################
ggplot(stroke_prediction, aes(x = age, y = stroke, color = work_type)) + geom_point() + labs(x = 'Age', y = 'Stroke') + ggtitle('Work Type by Age')
#########################################################################################

```

Correlation Analysis between numeric variables
```{r}
# Change the int variables to numeric
hypertension   <- as.numeric(stroke.df$hypertension)
heart_disease  <- as.numeric(stroke.df$heart_disease)
stroke         <- as.numeric(stroke.df$stroke)
age            <- stroke.df$age
avg_glucose    <- stroke.df$avg_glucose_level

# Recoding some variables
gender <- as.numeric(recode(stroke.df$gender, "Female" = 0, "Male" = 1, "Other" = 2))

ever_married <- as.numeric(recode(stroke.df$ever_married, "No" = 0, "Yes" = 1))

work_type <- as.numeric(recode(stroke.df$work_type, "Private" = 0, "Self-employed" = 1, "Govt_job" = 2, "Never_worked" = 3, "children" = 4))

smoking_status <- as.numeric(recode(stroke.df$smoking_status, "formerly smoked" = 1, "never smoked" = 1, "smokes" = 2, "Unknown" = 3))

# Put all numeric vectors into a data frame
stroke_num <- data.frame(gender, age, smoking_status, ever_married, work_type, heart_disease, hypertension, avg_glucose, stroke)

# Create a correlation matrix
stroke_corr <- round(cor(stroke_num), 2)
melted_stroke <- melt(stroke_corr)

head(melted_stroke)
```

Correlation Plot
```{r echo=TRUE}
# Create the correlation plot
ggplot(data = melted_stroke, aes(x = X1, y = X2, fill = value, label = value)) + geom_tile() + scale_fill_gradient2(low = "#0C6291", high = "#A63446", mid = "#FBFEF9") + geom_text(aes(X2, X1, label = value), color = "black", size = 4) + labs(fill = "Pearson's\nCorrelation") + theme(axis.text.x = element_text(angle = 90)) + ggtitle("Correlations in Stroke Data")

```